---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}

## Loading the dataset
data<-read.csv("activity.csv")

## Converting the 5-min interval in time
interv<-as.character(format(as.numeric(levels(factor(data$interval)))/100,nsmall=2))
data$hour<-sub("\\.",":",interv)
data$time<-as.POSIXct(data$hour, format="%H:%M")
head(data)

```


## What is mean total number of steps taken per day?

```{r total_nb_step_per_day }
stepday<-aggregate(steps ~ date, data=data, sum)
hist(stepday$steps, xlab="Number of steps taken each day", main="Histogram of number of steps per day")

## Calculate mean and median and report them on histogram
stepday_mean<-mean(stepday$steps)
stepday_median<-median(stepday$steps)

abline(v=stepday_mean, col="red")
abline(v=stepday_median, col="magenta")
## Mean of the total number of steps per day
stepday_mean
## Median of the total number of steps per day
stepday_median
```
  
The mean of the total number of steps per day is **`r format(stepday_mean, nsmall=2)`** (red line on histogram)  
The median of the total number of steps per day is **`r stepday_median`** (magenta line on histogram)
  

## What is the average daily activity pattern?

```{r daily_activity_pattern }

stepinterv<-aggregate(steps ~ time+hour, data=data, mean)

plot(x=stepinterv$time,y=stepinterv$steps,type="l", main="Daily activity pattern", xlab="Time", ylab="Average number of steps")

```



## Imputing missing values

```{r}
## Nb of missing data per features
missingdata<-apply(data[,1:3],2,function(x) sum(is.na(x)))
missingdata

## Total nb of missing values
totalmissingdata<-sum(missingdata)
totalmissingdata
```
In the dataset, there are **`r totalmissingdata`** missing values  

### Creation of new datasets with estimated values for missing values

Missing data will be replaced by the mean at that time (5min interval)
```{r}
q<-merge(data,stepinterv,by.x="hour",by.y="hour")
indexna<-is.na(q$steps.x)
q2<-q
q2$steps.x[indexna]<-round(q$steps.y[indexna])
## Creation of new dataset (data2) with missing data replaced
data2<-data
data2$steps<-q2$steps.x
head(data2)
```
## Rebuilding the histogram of number of steps per day with new dataset

```{r total_nb_step_per_day2}
stepday2<-aggregate(steps ~ date, data=data2, sum)
hist(stepday2$steps, xlab="Number of steps taken each day", main="Histogram of number of steps per day [new dataset]")

## Calculate mean and median and report them on histogram
stepday_mean2<-mean(stepday2$steps)
stepday_median2<-median(stepday2$steps)

abline(v=stepday_mean2, col="red")
abline(v=stepday_median2, col="magenta")
## Mean of the total number of steps per day
stepday_mean2
## Median of the total number of steps per day
stepday_median2
```

The mean of the total number of steps per day is **`r format(stepday_mean2, nsmall=2)`** (red line on histogram)  
The median of the total number of steps per day is **`r format(stepday_median2,nsmall=2)`** (magenta line on histogram)  

```{r}
gap_mean<-round(100*abs(stepday_mean2-stepday_mean)/stepday_mean,2)
gap_mean
gap_median<-round(100*abs(stepday_median2-stepday_median)/stepday_median,2)
gap_median
```

Imputing missing data has a negligible impact on mean (`r format(gap_mean,nsmall=2)`%) but has a **significant impact on the median (`r format(gap_median,nsmall=2)`%)**  


## Are there differences in activity patterns between weekdays and weekends?

### Creation of 
```{r}
## Definition of weekday/weekend
dict_weekday<-data.frame("day"=weekdays(Sys.Date()+0:6),"type_day"=as.factor(c(rep("weekday",5),rep("weekend",2))))

## Creation of new column weekday/weekend (and of a temporary technical column : day)
data2$day<-weekdays(as.Date(data2$date))
data2<-merge(data2,dict_weekday,by.x="day",by.y="day")
## deletion of technical feature
data2$day<-NULL

head(data2)

## 2-facets plot
library(lattice)
stepinterv2<-aggregate(steps ~ interval+type_day, data=data2, mean)
#stepinterv2<-aggregate(steps ~ time, data=data2, mean)
#plot(x=stepinterv2$time,y=stepinterv2$steps,type="l", main="Daily activity pattern", xlab="Time", ylab="Average number of steps")
xyplot(steps~interval | type_day,data=data2,type="l")

```

